International impact of large multi-centre surgical trials of arthroscopic subacromial decompression

### 2021-11-09 Paper-ASAD-rates.Rmd
Minor edits and github link

@@

Observational retrospective study.

@@

* This is most comprehensive study of ASAD incidence we are aware of. Routinely collected datasets were used to assess proportions of the patients undergoing this procedure in several countries
* Standardised case definitions were used across databases to increase comparability of findings
* Temporal changes in database coverage and quality of reporting can influence findings.  The observed variation in ASAD incidence may not be entirely attributable to changes in ASAD surgery rates.

# Introduction 
Evidence-based medicine requires that high-quality evidence be translated into widespread clinical practice. Randomised surgical trials are increasing in frequency, yet their influence on changing surgical practice is not well characterised. In this study, we investigate the changing incidence of the most common shoulder operation (arthroscopic subacromial decompression) before, during and after the conduct of two large pragmatic randomised controlled trials[@beardArthroscopicSubacromialDecompression2018; @paavolaSubacromialDecompressionDiagnostic2018] and a rapid recommendation[@vandvikSubacromialDecompressionSurgery2019].

@@

Anecdotal evidence of the international response in the provision of ASAD of different health care systems to these high-quality RCT findings seems to suggest that the incorporation of research evidence into practice is varied. In the UK and within 6 months of publication of the CSAW trial, NHS England put ASAD on a list of ‘procedures of limited value’, meaning that it should only be performed in limited circumstances[@EvidenceBasedInterventionsConsultation2018]. In Scotland, ASAD declined consistently from 2014-2018, with the largest decrease from 2017 to 2018 coinciding with the publication of the two randomised trials[@jenkinsChangingIncidenceArthroscopic2020]. However, data evaluating changes in ASAD incidence since publication of the trials in the rest of the UK are not yet available.

@@

## Study design and setting 
We did a retrospective study to describe ASAD incidence in Europe and the US and examined the trends in characteristics of patients undergoing ASAD. Incidence data were stratified according to age and gender, and patient characteristics were age, gender, Charlson index [@charlsonNewMethodClassifying1987] and medicine use in the previous month (additional characteristics were examined and are given in the aggregated data on GitHub where feasible: XXX[link to be added]).

@@

## Participants
Patients were eligible for inclusion in the cohort if they had at least one visit recorded in the database in a given year and cases were defined as patients undergoing ASAD for the first time in their records in a given year. The following codes were used to identify ASAD cases and were identified by two authors with surgical expertise (AJC and JR):

@@

Reasons for the lack of change in rates in the US are unclear. It is possible that the “ongoing RCT effect” is specific to countries or populations in which trials are conducted, and that results of surgical RCTs take more time to change practice internationally than nationally, despite the international dissemination of results. However, data from Sweden suggest a decrease in ASAD similar to that observed in our study, and the results of the trials are rapidly informing clinical guidelines (personal communication, Stefan Lohmander). Clinicians in countries such as the US may not view the results of the studies conducted in UK and Finland to be generalisable to their practice; however, the patient cohorts recruited to these trials are clearly defined in the trial publications and are not unique to the country of origin.

@@

Certain codes used to identify the cohort may result in the inclusion of cases that are not arthroscopic (i.e. are open) SAD or not ASAD specifically for shoulder impingement. Acromioplasty or acromionectomy, partial, with or without coracoacromial ligament release, for example, could potentially represent open SAD. However, open SAD is much less common, for example in the US comprising on 5% of SAD surgeries performed[@curtisNationalTrendsShow2021]. ASADs performed as part of other shoulder operations like rotator cuff repair or resection of calcific deposit may not have been excluded with the used codes. There is inevitably a trade-off between sensitivity and specificity of codes; with two experienced shoulder surgeons helping in our cohort definition, we feel the codes used were reasonable.

@@

# Reproducibility and data availability
This article was written in RMarkdown and can be reproduced from the publicly available aggregate data and code provided on GitHub. Some data had to be removed because it contained cells with low counts which could compromise anonymity. RStudio[@rstudioteamRStudioIntegratedDevelopment2020] and several R[@rcoreteamLanguageEnvironmentStatistical2020a] packages[@kassambaraGgpubrGgplot2Based2020; @usheyRenvProjectEnvironments2020; @wickhamReadxlReadExcel2019; @wickhamScalesScaleFunctions2020; @wickhamWelcomeTidyverse2019; @xieKnitrGeneralpurposePackage2021; @zhuKableExtraConstructComplex2020] were used.

>>>>

\newpage

@@

Observational study of incidence rates.

@@

* This is the most comprehensive study of ASAD incidence we are aware of. Routinely collected datasets were used to assess proportions of the patients undergoing this procedure in several countries
* Standardised case definitions were used across databases to increase comparability of findings
* Temporal changes in database coverage and quality of reporting can influence findings.  The observed variation in ASAD incidence may not be entirely attributable to changes in ASAD surgery rates.

\newpage

# Introduction 
Evidence-based medicine requires that high-quality evidence be translated into widespread clinical practice. Randomised surgical trials are increasing in frequency, yet their influence on changing surgical practice is not well characterised. In this study, we investigate the changing incidence of the most common shoulder operation (arthroscopic subacromial decompression) before, during and after the conduct of two large pragmatic randomised controlled trials[@beardArthroscopicSubacromialDecompression2018; @paavolaSubacromialDecompressionDiagnostic2018] and publication of a rapid recommendation[@vandvikSubacromialDecompressionSurgery2019].

@@

Anecdotal evidence of the international response in the provision of ASAD of different health care systems to these RCT findings seems to suggest that the incorporation of research evidence into practice is varied. In the UK and within 6 months of publication of the CSAW trial, NHS England put ASAD on a list of ‘procedures of limited value’, meaning that it should only be performed in limited circumstances[@EvidenceBasedInterventionsConsultation2018]. In Scotland, ASAD declined consistently from 2014-2018, with the largest decrease from 2017 to 2018 coinciding with the publication of the two randomised trials[@jenkinsChangingIncidenceArthroscopic2020]. However, data evaluating changes in ASAD incidence since publication of the trials in the rest of the UK are not yet available.

@@

## Study design and setting 
We did a retrospective study to describe ASAD incidence in Europe and the US and examined the trends in characteristics of patients undergoing ASAD. Incidence data were stratified according to age and gender, and patient characteristics were age, gender, Charlson index [@charlsonNewMethodClassifying1987] and medicine use in the previous month (additional characteristics were examined and are given in the aggregated data on GitHub, see "Reproducibility and data availability").

@@

## Participants
Patients were eligible for inclusion in the cohort if they had at least one visit recorded in the database in a given year and cases were defined as patients undergoing ASAD for the first time in their records in a given year. The following codes were used to identify ASAD cases and were identified by two authors with surgical expertise (AJC and JLR):

@@

Reasons for the lack of change in rates in the US are unclear. It is possible that any “ongoing RCT effect” is specific to countries or populations in which trials are conducted, and that results of surgical RCTs take more time to change practice internationally than nationally, despite the international dissemination of results. However, data from Sweden suggest a decrease in ASAD similar to that observed in our study, and the results of the trials are rapidly informing clinical guidelines (personal communication, Stefan Lohmander). Clinicians in countries such as the US may not view the results of the studies conducted in UK and Finland to be generalisable to their practice; however, the patient cohorts recruited to these trials are clearly defined in the trial publications and are not unique to the country of origin.

@@

Certain codes used to identify the cohort may result in the inclusion of cases that are not arthroscopic (i.e. are open) SAD or not ASAD specifically for shoulder impingement. Acromioplasty or acromionectomy, partial, with or without coracoacromial ligament release, for example, could potentially represent open SAD. However, open SAD is much less common, for example in the US comprising only 5% of SAD surgeries performed[@curtisNationalTrendsShow2021]. ASADs performed as part of other shoulder operations like rotator cuff repair or resection of calcific deposit may not have been excluded with the used codes. There is inevitably a trade-off between sensitivity and specificity of codes; with two experienced shoulder surgeons helping in our cohort definition, we feel the codes used were reasonable.

@@

# Reproducibility and data availability
This article was written in RMarkdown and can be reproduced from the publicly available aggregate data and code provided on GitHub (https://github.com/worcjamessmith/ASAD_rates_public). Some data could not be shared because it contained cells with low counts which could compromise anonymity. RStudio[@rstudioteamRStudioIntegratedDevelopment2020] and several R[@rcoreteamLanguageEnvironmentStatistical2020a] packages[@kassambaraGgpubrGgplot2Based2020; @usheyRenvProjectEnvironments2020; @wickhamReadxlReadExcel2019; @wickhamScalesScaleFunctions2020; @wickhamWelcomeTidyverse2019; @xieKnitrGeneralpurposePackage2021; @zhuKableExtraConstructComplex2020] were used.
